Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Remove scaling of SPF

\section{Introduction}
    We will in this project study the Variational Monte Carlo method, and use it
    to evaluate the ground state energy of a trapped, hard sphere Bose gas.
    \cite{dubois2001bose} \cite{nilsen2005vortices}

@@

        \subsubsection{Scaling the system}
            We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
            \cite{dubois2001bose}, where
            \begin{align}
                a_{\text{ho}}
                &=
                \sqrt{\frac{\hbar}{m\omega}}.
            \end{align}
            We can then rewrite the Hamiltonian for the elliptic potential in terms
            of this new distance. By doing a variable substitution for each
            direction in the Laplace operator we get
            \begin{align}
                \nabla_k^2
                &=
                \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
            \end{align}
            The one-body part of the Hamiltonian then yields
            \begin{align}
                H
                &=
                \sum_{k = 1}^N
                \brac{
                    -\frac{\hbar^2}{2m}\nabla_k^2
                    + \half m\brak{
                        \omega^2(x_k^2 + y_k^2)
                        + \omega_z^2 z_k^2
                    }
                }
                \\
                &=
                \frac{\hbar\omega}{2}
                \sum_{k = 1}^N
                \brac{
                    -{\nabla_k'}^2
                    +
                    \brak{
                        {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
                    }
                },
                \label{eq:scaled_hamiltonian}
            \end{align}
            where we have introduced the dimensionless frequency $\lambda =
            \omega_z/\omega$. The single particle functions also get a scaling
            factor
            \begin{align}
                \phi_k
                &=
                \exp\brak{
                    -\frac{\alpha\hbar}{m\omega}
                    ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
                }.
                \label{eq:scaled_spf}
            \end{align}
            The correlation wavefunction and the interaction potential remain the
            same, but using the scaled hard core radius $a/a_{\text{ho}} =
            0.0043$.

@@

>>>>

\section{Introduction}
    We will in this project study the Variational Monte Carlo method, and use it
    to evaluate the ground state energy of a trapped, hard sphere Bose gas.

@@

        \subsubsection{Scaling the system}
            We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
            \cite{dubois2001bose}, where
            \begin{align}
                a_{\text{ho}}
                &=
                \sqrt{\frac{\hbar}{m\omega}}.
            \end{align}
            We can then rewrite the Hamiltonian for the elliptic potential in terms
            of this new distance. By doing a variable substitution for each
            direction in the Laplace operator we get
            \begin{align}
                \nabla_k^2
                &=
                \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
            \end{align}
            The one-body part of the Hamiltonian then yields
            \begin{align}
                H
                &=
                \sum_{k = 1}^N
                \brac{
                    -\frac{\hbar^2}{2m}\nabla_k^2
                    + \half m\brak{
                        \omega^2(x_k^2 + y_k^2)
                        + \omega_z^2 z_k^2
                    }
                }
                \\
                &=
                \frac{\hbar\omega}{2}
                \sum_{k = 1}^N
                \brac{
                    -{\nabla_k'}^2
                    +
                    \brak{
                        {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
                    }
                },
                \label{eq:scaled_hamiltonian}
            \end{align}
            where we have introduced the dimensionless frequency $\lambda =
            \omega_z/\omega$. The wavefunction does not need to be scaled as the
            exponentials are already dimensionless. We only need to use the
            scaled hard core radius $a' = a/a_{\text{ho}}$. We are interested in
            using the radius $a' = 0.0043$ as in the references
            \cite{dubois2001bose} and \cite{nilsen2005vortices}.

@@

                This is a so-called "hacky" solution, but it does the trick. It
                is also the reason why the result figure in
                \autoref{fig:gradient_descent_interacting}.
